---
title: Cleaning the threat dataset
author: Gareth Kindler
output:
    html_document:
        code_folding: show
---

This document details the cleaning of the threat data for to the Threatened Australians (threatened.org.au) project.

## Libraries

```{r, eval = FALSE, class.source = 'fold-hide'}
library(tidyverse)
library(sf)
library(magrittr)
library(jsonlite)
library(units)
library(readxl)
library(stringr)
library(rmapshaper)
library(httpgd)
```

## Import: Threatened Species

### Data key

`threats`:
For the most up-to-date information on threats facing Australian imperilled species:

>  Ward, M., Carwardine, J., Yong, C. J., Watson, J. E. M., Silcock, J., Taylor, G. S., Lintermans, M., Gillespie, G. R., Garnett, S. T., Woinarski, J., Tingley, R., Fensham, R. J., Hoskin, C. J., Hines, H. B., Roberts, J. D., Kennard, M. J., Harvey, M. S., Chapple, D. G., & Reside, A. E. (2021). A national-scale dataset for threats impacting Australia’s imperiled flora and fauna. Ecology and Evolution, 11, 11749– 11761. https://doi.org/10.1002/ece3.7920

```{r}
threats <- readxl::read_excel(
    "data/Ward_2019_national_dataset_imperiled_appendix_S1.xlsx",
    "Species-Threat-Impact"
)
species <- st_read(
    "output/clean_data/species_clean.gpkg"
)
```

## Clean: Threats

As the project has no target audience and wants to be accessible for all persons, we opted to focus on broad scale threats identified through the IUCN criteria used in the research article. And this has added benefits of simplicity to the project.

We then recoded the data with a `threat_ID` for linking attributes in our SQL database. And removed species not in common with the species data.

```{r}
names(threats) <- make.names(names(threats), unique = TRUE)

threats_clean <- threats %>%
    rename(
        species_name_adjusted = Species.name.adjusted,
        scientific_name = Species.name,
        vernacular_name = Common.name,
        broad_level_threat = Broad.level.threat,
        taxon_group = Group
    ) %>%
    select(
        species_name_adjusted, scientific_name,
        vernacular_name, broad_level_threat,
        taxon_group
    ) %>%
    group_by(
        species_name_adjusted, broad_level_threat
    ) %>%
    summarise() %>%
    ungroup() %>%
    rename(
        scientific_name = species_name_adjusted
    ) %>%
    inner_join(species) %>%
    select(
        taxon_ID, broad_level_threat
    ) %>%
    mutate(
        threat_ID =
            case_when(
                broad_level_threat == "Adverse fire regimes" ~ "T01",
                broad_level_threat == "Changed surface and groundwater regimes" ~ "T02",
                broad_level_threat == "Climate change and severe weather" ~ "T03",
                broad_level_threat == "Disrupted ecosystem and population processes" ~ "T04",
                broad_level_threat == "Habitat loss, fragmentation and degradation" ~ "T05",
                broad_level_threat == "Invasive species and diseases" ~ "T06",
                broad_level_threat == "Overexploitation and other direct harm from human activities" ~ "T07",
                broad_level_threat == "Pollution" ~ "T08"
            )
    ) %T>%
    # mutate(
    #     broad_level_threat_alt =
    #         case_when(
    #             broad_level_threat == "Adverse fire regimes" ~ "Adverse fire regimes",
    #             broad_level_threat == "Changed surface and groundwater regimes" ~ "Changed surface and groundwater regimes",
    #             broad_level_threat == "Climate change and severe weather" ~ "Climate change and severe weather",
    #             broad_level_threat == "Disrupted ecosystem and population processes" ~ "Disrupted ecosystem and population processes",
    #             broad_level_threat == "Habitat destruction, fragmentation and degradation" ~ "Habitat loss, fragmentation and degradation",
    #             broad_level_threat == "Invasive species and diseases" ~ "Invasive species and diseases",
    #             broad_level_threat == "Overexploitation and other direct harm from human activities" ~ "Overexploitation and other direct harm from human activities",
    #             broad_level_threat == "Pollution" ~ "Pollution"
    #         )
    write_json(
        "output/clean_data/threats_clean.json"
    )
```

### Collapse: Threats

```{r}
threats_collapsed_clean <- threats_clean %>%
    group_by(
        taxon_ID
    ) %>%
    summarise(
        threat_ID_collapsed = paste(threat_ID, collapse = ", ")
    ) %>%
    ungroup() %T>%
    write_json(
        "output/clean_data/threats_collapsed_clean.json"
    )
```
