---
title: Compiling the clean and analysed data into tables
author: Gareth Kindler
output:
    html_document:
        code_folding: show
---

This document details the compiling of data into tables for the database of the Threatened Australians (threatened.org.au) project. It is the final of four documents.

## Libraries

```{r, eval = FALSE, class.source = 'fold-hide'}
library(tidyverse)
library(sf)
library(rmapshaper) # For installing use 'library(remotes)'
library(jsonlite)
library(magrittr)
library(units)
library(httpgd)
```

## Import: Electoral

Explanations of what each of these data represent can be found in `2.local_attributes_clean.r`

```{r, eval = FALSE}
elects <- st_read(
    "output/analysed_data/final/elects_tbl.gpkg"
)
postcodes_elects <- st_read(
    "output/analysed_data/final/postcodes_elects_tbl.gpkg"
)
MP_info <- fromJSON(
    "output/clean_data/MP_info_clean.json"
)
MP_voting_info <- fromJSON(
    "output/clean_data/MP_voting_info_clean.json"
)
demo <- fromJSON(
    "output/clean_data/demo_clean.json"
)
```

## Import: Threatened Species

Explanations of what each of these data represent can be found in `2.local_attributes_clean.r`

```{r, eval = FALSE}
species_elects <- st_read(
    "output/analysed_data/final/species_elects_tbl.gpkg"
)
species <- st_read(
    "output/analysed_data/final/species_tbl.gpkg"
)
threats <- fromJSON(
    "output/clean_data/threats_clean.json"
)
threats_collapsed <- fromJSON(
    "output/clean_data/threats_collapsed_clean.json"
)
animals <- fromJSON(
    "output/clean_data/species_FT_animals_clean.json"
) %>%
    select(taxon_ID)
plants <- fromJSON(
    "output/clean_data/species_plants_clean.json"
) %>%
    select(taxon_ID)
animals_images <- fromJSON(
    "output/clean_data/animals_images_clean.json"
) %>%
    mutate(
        taxon_ID = as.character(taxon_ID)
    )
animals_info <- fromJSON(
    "output/clean_data/animals_info_clean.json"
) %>%
    mutate(
        taxon_ID = as.character(taxon_ID)
    )
```

## Import: Other

`action_groups` is a manually created table that details the environmental NGOs we want to highlight on the website.

```{r, eval = FALSE}
action_groups <- fromJSON(
    "raw_data/action_groups.json"
)
```

## Filter: Animals

This is previously referred to as the filter dataframe (`2.local_attributes_clean.Rmd`) we use to remove species that aren't relevant to the project such as species that don't intersect with any federal electorates. Or that we don't have essential information on such as threats facing them.

```{r, eval = FALSE}
species_ft <- species_elects %>%
    st_set_geometry(NULL) %>%
    group_by(taxon_ID) %>%
    summarise() %>%
    inner_join(threats_collapsed) %>%
    select(taxon_ID) %>%
    mutate(taxon_ID = as.character(taxon_ID)) %T>%
    write_json(
        "output/clean_data/species_ft.json"
    )
```

## Table: Species-elects

The bulk of the following code chunk is dedicated to creating an `species_electorate_coverage` attribute. This is a string that we can showcase on each of the unique combinations of species-elects that occur (~4K). The string describes how unique the species is to the electorate(s).

```{r, eval = FALSE}
animals_elects_tbl <- species_elects %>%
    mutate(
        across(
            c(
                species_range_area_sqkm,
                species_intersect_area_sqkm,
                percent_range_within
            ),
            signif,
            digits = 3
        )
    ) %>%
    group_by(
        taxon_ID
    ) %>%
    mutate(species_range_intersects_with_n_electorates = n_distinct(electorate)) %>%
    ungroup() %>%
    mutate(
        scientific_name_clean = word(
            scientific_name, 1, 1,
            sep = fixed("(")
        )
    ) %>%
    # filter(
    #     percent_range_within >= 0.05 # TODO: gotta solve this among others
    # )
    # This introduces various errors, but for most species it's going to be good
    mutate(
        species_electorate_coverage = case_when(
            percent_range_within == 1 ~ paste0(
                scientific_name_clean,
                " is only found within ",
                electorate
            ),
            percent_range_within >= 0.8 &
            percent_range_within < 1
            # species_range_intersects_with_n_electorates != 1
             ~ paste0(
                scientific_name_clean,
                " has greater than 80% of it's range within ",
                electorate
            ),
            TRUE ~ paste0(
                scientific_name_clean,
                " is found across ",
                species_range_intersects_with_n_electorates,
                " electorates."
            )
        )
    ) %>%
    select(
        taxon_ID, scientific_name_clean,
        species_range_intersects_with_n_electorates,
        species_range_area_sqkm, species_electorate_coverage, electorate,
        species_intersect_area_sqkm, percent_range_within, geom
    ) %>%
    inner_join(species_ft) %>%
    inner_join(animals) %T>%
    st_write(
        "output/analysed_data/final/imports/animals_elects_tbl.geojson",
        layer = "animals_elects_tbl", append = FALSE, delete_dsn = TRUE
    )
```

## Table: Plants-elects

Two `vernacular_name`s of plants were missing indefinite articles which were creating some grammatical errors.

```{r, eval = FALSE}
plants_elects_tbl <- species_elects %>%
    st_set_geometry(NULL) %>%
    inner_join(plants) %>%
    mutate(
        vernacular_name_first_clean = replace(
            vernacular_name_first_clean,
            vernacular_name_first_clean == "cycad",
            "a cycad "
        )
    ) %>%
    mutate(
        vernacular_name_first_clean = replace(
            vernacular_name_first_clean,
            vernacular_name_first_clean == "fern",
            "a fern"
        )
    ) %>%
    select(
        taxon_ID, scientific_name_clean, electorate,
        vernacular_name_first_clean, threatened_status,
        SPRAT_profile
    ) %>%
    relocate(electorate, .after = SPRAT_profile) %T>%
    write_json(
        "output/analysed_data/final/imports/plants_elects_tbl.json"
    )
```

## Table: Animals

```{r, eval = FALSE}
animals_tbl <- species %>%
    inner_join(animals) %>%
    inner_join(species_ft) %>%
    inner_join(animals_info) %>%
    inner_join(animals_images) %>%
    inner_join(threats_collapsed) %>%
    mutate(
        vernacular_name_first_clean = replace(
            vernacular_name_first_clean,
            vernacular_name_first_clean == "freshwater crayfish",
            "a freshwater crayfish"
        )
    ) %>%
    mutate(
        vernacular_name_first_clean = replace(
            vernacular_name_first_clean,
            vernacular_name_first_clean == "A native bee",
            "a native bee"
        )
    ) %>%
    relocate(
        vernacular_name_first,
        .after = vernacular_name
    ) %>%
    relocate(
        vernacular_name_other,
        .after = vernacular_name_first
    ) %>%
    relocate(
        vernacular_name_first_clean,
        .after = vernacular_name_first
    ) %>%
    select(
         taxon_ID, scientific_name_clean, vernacular_name_first_clean,
         threatened_status, migratory_status, taxon_group,
         taxon_kingdom, SPRAT_profile,
         species_range_area_sqkm, description, ALA_URL, image_URL,
         threat_ID_collapsed, geom
    ) %>%
    mutate(
        across(
            species_range_area_sqkm,
            signif,
            digits = 3
        )
    ) %T>%
    st_write(
        "output/analysed_data/final/imports/animals_tbl.geojson",
        layer = "animals_tbl", append = FALSE, delete_dsn = TRUE
    )
```

## Table: Elects

On the *Browser* page, there is the option to explore other electorates by a couple of filters, one such is by the number of animals and plants found within the boundaries of each electorate.

```{r, eval = FALSE}
animals_elects_counts <- species_elects %>%
    st_set_geometry(NULL) %>%
    inner_join(animals) %>%
    inner_join(species_ft) %>%
    group_by(electorate) %>%
    summarise(no_animal_species = n_distinct(taxon_ID))
plants_elects_counts <- species_elects %>%
    st_set_geometry(NULL) %>%
    inner_join(plants) %>%
    inner_join(species_ft) %>%
    group_by(electorate) %>%
    summarise(no_plant_species = n_distinct(taxon_ID))
```

And then compiling all the electoral information into a single `elects_tbl`.

```{r, eval = FALSE}
elects_tbl <- elects %>%
    inner_join(demo) %>%
    inner_join(MP_info) %>%
    inner_join(MP_voting_info) %>%
    inner_join(animals_elects_counts) %>%
    full_join(plants_elects_counts) %>%
    mutate(
        across(
            electorate_area_sqkm,
            signif,
            digits = 3
        )
    ) %T>%
    st_write(
        "output/analysed_data/final/imports/elects_tbl.geojson",
        layer = "elects_tbl", append = FALSE, delete_dsn = TRUE
    )
```

## Table: Postcodes-elects

As alluded to in `3.HPC_spatial_ops.Rmd`, intersecting `postcodes` and `elects` will cause intersections to be produced where there is slight overlaps, to an excessive degree. The slight overlaps are miniscule and often too small to be residential areas. We dealt with this by filtering out at the proportion of POA found within each electorate when it is less or equal to `0.001`.

```{r, eval = FALSE}
postcodes_elects_tbl <- postcodes_elects %>%
    st_set_geometry(NULL) %>%
    mutate(
        across(
            c(
                POA_area_sqkm, POA_elect_int_area_sqkm
            ),
            signif,
            digits = 3
        )
    ) %>%
    mutate(
        proportion_POA_within_elect = POA_elect_int_area_sqkm / POA_area_sqkm
    ) %>%
    filter(
        !proportion_POA_within_elect <= 0.001
    ) %T>%
    write_json(
        "output/analysed_data/final/imports/postcodes_elects_tbl.json"
    )
```

## Table: Action-groups

`Action_groups` just needs to be passed into the data import folder.

```{r, eval = FALSE}
action_groups_tbl <- action_groups %T>%
    write_json(
        "output/analysed_data/final/imports/action_groups_tbl.json"
    )
```

## Table: Threats

As the `species_tbl` has the `threat_ID` for the broad level threats each species faces, we just need a linking table for more information.

```{r, eval = FALSE}
threats_tbl <- threats %>%
    group_by(
        broad_level_threat, threat_ID
    ) %>%
    summarise() %>%
    relocate(
        threat_ID,
        .before = broad_level_threat
    ) %T>%
    write_json(
        "output/analysed_data/final/imports/threats_tbl.json"
    )
```
