---
title: Cleaning the threat dataset<br><h2>Threatened Australians (threatened.org.au)
author: Gareth Kindler<br>The University of Queensland, Australia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
        code_folding: show
---

```{r setup, include=FALSE, fig.align='center', warning=FALSE, message=FALSE}
# knitr::opts_chunk$set(echo = TRUE, comment = "#")
knitr::opts_knit$set(root.dir = "../")
```

## Libraries

```{r, class.source = 'fold-hide'}
library(tidyverse)
library(sf)
library(magrittr)
library(jsonlite)
library(units)
library(readxl)
library(stringr)
library(rmapshaper)
library(httpgd)
```

## Import: Threatened Species

### Data key

`threats`:
For the most up-to-date information on threats facing Australian imperilled species:

>  Ward, M., Carwardine, J., Yong, C. J., Watson, J. E. M., Silcock, J., Taylor, G. S., Lintermans, M., Gillespie, G. R., Garnett, S. T., Woinarski, J., Tingley, R., Fensham, R. J., Hoskin, C. J., Hines, H. B., Roberts, J. D., Kennard, M. J., Harvey, M. S., Chapple, D. G., & Reside, A. E. (2021). A national-scale dataset for threats impacting Australia’s imperiled flora and fauna. Ecology and Evolution, 11, 11749– 11761. https://doi.org/10.1002/ece3.7920

```{r}
threats <- readxl::read_excel(
    "data/Ward_2019_national_dataset_imperiled_appendix_S1.xlsx",
    "Species-Threat-Impact"
)
species <- st_read(
    "output/clean_data/species_clean.gpkg"
)
species_ft <- read_json(
    "output/clean_data/species_ft.json"
)
```

## Clean: Threats

As the project has no target audience and wants to be accessible for all persons, we opted to focus on broad scale threats identified through the IUCN criteria used in the research article. And this has added benefits of simplicity to the project.

We then recoded the data with a `threat_ID` for linking attributes in our SQL database. And removed species not in common with the species data.

```{r}
names(threats) <- make.names(names(threats), unique = TRUE)

threats_clean <- threats %>%
    rename(
        species_name_adjusted = Species.name.adjusted,
        scientific_name = Species.name,
        vernacular_name = Common.name,
        broad_level_threat = Broad.level.threat,
        taxon_group = Group
    ) %>%
    select(
        species_name_adjusted, scientific_name,
        vernacular_name, broad_level_threat,
        taxon_group
    ) %>%
    group_by(
        species_name_adjusted, broad_level_threat
    ) %>%
    summarise() %>%
    ungroup() %>%
    rename(
        scientific_name = species_name_adjusted
    ) %>%
    full_join(species) |>
    select(
        taxon_ID, broad_level_threat
    ) %>%
    mutate(
        broad_level_threat = replace_na(
            broad_level_threat,
            "No easily accessible threat information was found for this species. This does not mean there is no information out there. We encourage you to do a web search using the scientific latin name."
        )
    ) |>
    mutate(
        threat_ID =
            case_when(
                broad_level_threat == "Adverse fire regimes" ~ "T01",
                broad_level_threat == "Changed surface and groundwater regimes" ~ "T02",
                broad_level_threat == "Climate change and severe weather" ~ "T03",
                broad_level_threat == "Disrupted ecosystem and population processes" ~ "T04",
                broad_level_threat == "Habitat loss, fragmentation and degradation" ~ "T05",
                broad_level_threat == "Invasive species and diseases" ~ "T06",
                broad_level_threat == "Overexploitation and other direct harm from human activities" ~ "T07",
                broad_level_threat == "Pollution" ~ "T08",
                TRUE ~ "T09"
            )
    ) %T>%
    write_json(
        "output/clean_data/threats_clean.json"
    )
```

### Collapse: Threats

```{r}
threats_collapsed_clean <- threats_clean %>%
    group_by(
        taxon_ID
    ) %>%
    summarise(
        threat_ID_collapsed = paste(threat_ID, collapse = ", ")
    ) %>%
    ungroup() %T>%
    write_json(
        "output/clean_data/threats_collapsed_clean.json"
    )
```
